---
title: "DS-400_Penguins_Activity"
format: html
editor: visual
---

## Penguins Activity

Tallen Vidal

09/22/2026

## Model Penguin Data using Bayesian Regression

## 1) Loading Libraries

```{r}
options(scipen = 99)
library(tidyverse)
library(bayesrules)
library(rstanarm)
library(vembedr)
library(ggplot2)
library(ggpubr)
library(plotly)
library(report)
library(ggExtra)

data <- bayesrules::penguins_bayes
```

## 2) Plots for EDA

Two quick plots, one sentence each describing what you see:

- A histogram of `flipper_length_mm`. Where is it centered? Roughly bell-shaped, or doing something unusual?

- A scatterplot of `flipper_length_mm` (y) against `bill_length_mm` (x). Linear? Positive or negative? Strong or weak?

```{r}
### Histogram ###

ggplot(data = data, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram() +
  labs(title = "Flipper Length Distribution", x = "Flipper Length in mm", y = "Count") +
  theme_minimal()
```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
### Scatterplot ###

p<-ggplot(data, aes(x = bill_length_mm, y = flipper_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()
ggMarginal(p, type = "density")
```

```{r}
m0 <- 200   # centered intercept mean
s0 <-  10.20  # centered intercept SD
m1 <-   3.5 # slope mean
s1 <- 1.02 # slope SD
l  <- 0.08   # sigma's exponential rate
```

```{r}
plot_normal(200, 10.2)
```

```{r}
model <- stan_glm(
  flipper_length_mm ~ bill_length_mm, data = data,
  family = gaussian,
  prior_intercept = normal(m0, s0),
  prior = normal(m1, s1),
  prior_aux = exponential(l),
  chains = 4, iter = 2000, seed = 84735
)
```

```{r}
summary(model)
```

```{r}
plot_normal(1.7, 0.051)
```

```{r}
report::report(model)
```

## In Class Search Activity

Link to competitive swimming analysis using Bayesian allometric quantile regressions:

<https://pubmed.ncbi.nlm.nih.gov/32772638/>
