---
title: "DS-400_Penguins_Activity"
format: html
editor: visual
---

## Penguins Activity

Tallen Vidal

09/22/2026

## Model Penguin Data using Bayesian Regression

## 1) Loading Libraries

```{r}
options(scipen = 99)
library(tidyverse)
library(bayesrules)
library(rstanarm)
library(vembedr)
library(ggplot2)
library(ggpubr)
library(plotly)
library(report)
library(ggExtra)

data <- bayesrules::penguins_bayes
```

## 2) Plots for EDA

Two quick plots, one sentence each describing what you see:

- A histogram of `flipper_length_mm`. Where is it centered? Roughly bell-shaped, or doing something unusual?

- A scatterplot of `flipper_length_mm` (y) against `bill_length_mm` (x). Linear? Positive or negative? Strong or weak?

```{r}
### Histogram ###

ggplot(data = data, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram() +
  labs(title = "Flipper Length Distribution", x = "Flipper Length in mm", y = "Count") +
  theme_minimal()
```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
### Scatterplot ###

p<-ggplot(data, aes(x = bill_length_mm, y = flipper_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()
ggMarginal(p, type = "density")
```
